How explicit is the Explicit Formula?

### 2016-01-24 paper/paper.tex
move plots to subdir; rewrite intro

{\bf William:}  We need a sentence or paragraph saying what relationship our paper has to Riemann's work.

 Any `Explicit Formula' in analytic number theory deals with an {\it arithmetically interesting quantity}, often given as a partial sum $F(X)\cdot\sum_{p<X}G(p)$ of locally defined quantities $G(p)$ attached to prime numbers $p$, summed to up to some cutoff value, $p<X$  (and normalized for convenience by an elementary factor $F(X)$). The formula expresses this function of $X$  as a {\it dominant term}, plus an easily controlled error term, plus an interesting third term that might be called the {\it oscillatory term} that, in most cases, is only conjecturally controllable.

@@

  Specifically we will  focus on {\it issues of bias}  following the classical `Explicit Formula," and the work of:
Sarnak, \ Granville,\  Rubenstein,\, Martin-Watkins,\ Fiorilli,\ Bober,\ Conrey-Snaith, and others.
 The example-problem we consider is related to the question---given an elliptic curve over the rational numbers and letting  $p$ range through prime numbers---of how often  $p+1$ is an {\it over-count} or an {\it under-count} for the number of rational points on the curve modulo $p$? The rough answer is 50/50, but there can be a `bias' similar to the classical Chebyshev bias  \cite{R-S}. For such finer statistics one resorts to `Explicit Formulas.'  Here, computation can even outstrip theory in that people have algorithms to make such computations whether or not the holomorphicity of the $L$-functions in question have been proved. Computations have depended on the work of many people, notably Mike Rubenstein's ****

 We have used these notes to provide material for {\it unusually short lectures}. Specifically, the notes were initially written for each of our 20 minute lectures at the San Diego Conference, as we have already mentioned, but also one of us replayed it all{\footnote{ Well, not quite all$\dots$}} in a 6 minute and 40 second lecture at Harvard (See PechaKucha web page ****). We say this to make it clear that this article is meant to be  somewhat light reading; it contains no new results, as we have mentioned, and no proofs of anything. But we feel that the issues here are a good source of student projects, and that our text might be  a companion to any reader of the beautiful expository articles on classical Chebyshev bias, such as \cite{R-S}, or on the biases we discuss, as in \cite{S}.

@@

 Happily, the other of us (W.S.) has produced relevant computations that do exactly that (for applications of the Explicit Formula to certain problems of current interest to both of us). This, then, is a phenomenological talk, with (at least the beginning of)  a corpus  of graphs that offer some  illustration of the effectiveness---or non-effectiveness---of the explicit formula in the specific instances of interest to us.

@@

  \section{ The RHS of our Explicit Formulas}
   Here are some brief comments on each of the `terms' on the RHS of the Explicit  Formula for our three variants, where we write that RHS of--for example--the well-done variant above for an elliptic curve $E$  as given below:

@@

  \begin{conjecture}  $ S_E(X)\ \  {\stackrel{?}{=}}\ \  o(\log X).$\end{conjecture}   The analogous oscillatory term for the classical Riemann zeta function  has an extensive literature. See, for example \cite{G}, \cite{Fu} and the bibliography there. Here are some pictures to convey a sense of how our $S_E(X,T)$ behaves, at least in the currently computable range which (roughly) allows $T$ to be only as high as  $10^4$.
  \newpage
  \centerline{ $E = 11a$}
 \vskip10pt
  \includegraphics[width=0.9\textwidth]{OSC11.pdf}

  \centerline{ $E = 37a$}
  \includegraphics[width=0.9\textwidth]{OSC37.pdf}
%\vskip20pt
 % \includegraphics[width=1.0\textwidth]{OSC.pdf}
   \vskip20pt
  \centerline{ $E = 389a$}
\vskip20pt
  \includegraphics[width=0.9\textwidth]{OSC389.pdf}

@@

Let ${\R}_{>0}$ be the multiplicative group of positive real numbers, and ${\R}$ the additive group of reals. For $I \subset {\R}_{>0}$ a Haar measurable set, let $|I|$ denote a Haar measure.  Let ${\S}:{\R}_{>0} \to {\R}$ be a real-valued Lebesgue-integrable function.  Fixing $I \subset {\R}_{>0}$ a subset  of finite measure,  for every measurable subset $J\subset {\R}$, form the probability measure on ${\R}$
$$J\mapsto \mu_{{\S},I}(J): = {\frac{|I\cap {\S}^{-1}(J)|}{|I|}}.$$  So, $\mu_{{\S},I}(J)$ is the {\it probability that the function ${\S}$ achieves a value in the range $J$ over the gamut of arguments in $I$.}   Say that ${\S}$ has a {\bf normal distribution of values} if, for $X > 0$ setting $I_X= (0,X]$, the limit  $$\mu_{{\S}}:= \lim_{X \to \infty}\mu_{{\S},I_X}$$ exists. These  definitions are particularly relevant to the oscillatory terms ${\S}(X):= S_E(X)$ that we are currently studying. The data seems to indicate convergence to a limiting distribution (the  mean  value being $0$) with a strikingly small (variance, or equivalently: strikingly small) standard deviation of values.
\vskip20pt
Here, then, are some pictures  of what seems to be data 'converging' to a limiting  distribution $\mu_E$ of the values of the oscillatory terms $S_E(X)$ for a few elliptic curves $E$:
  \vskip10pt
  \centerline{ $E = 11a$}
 \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{bite11.pdf}
    \vskip10pt
  \centerline{ $E = 37a$}
  \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{bite37.pdf}
    \vskip10pt
       The red curve is the normal distribution with mean $0$ and standard deviation given by that of the data.
   \vskip10pt
    {\bf  Note: } Conditional on the conjecture $LI(E)$ (see section \ref{Fi} below) $\mu_E$ exists (see \ref{S}).

     It is interesting  to compare $\mu_E$ to the limiting distributions connected to the bias of nonresidues to residues mod $q$, as in \cite{R-S}. There one has the added feature that these limiting distributions themselves tend to the normal distribution as the modulus q tends to infinity.
   % \includegraphics[width=0.9\textwidth]{bite389.pdf}
 %\vskip30pt
  \centerline{ $E = 389a$}
  \vskip10pt
\hskip100pt  \includegraphics[width=0.6\textwidth]{bite389.pdf}

@@

 \centerline{ $E = 11a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{11a40t60.pdf}
    \newpage
  \centerline{ $E = 11a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{11a60t100.pdf}

   \vskip10pt
   \centerline{ $E = 11a$  between $990$ and $1010$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{11a990t1010.pdf}

@@

   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{37a40t60.pdf}

   \centerline{ $E = 37a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{37a60t100.pdf}
 \newpage
  \centerline{ $E = 37a$  between $990$ and $1010$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{37a990t1010.pdf}
 \newpage
  \centerline{ $E = 389a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{389a40t60.pdf}

   \centerline{ $E = 389a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{389a60t100.pdf}
\newpage
  \centerline{ $E = 5077a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{5077a40t60.pdf}

   \centerline{ $E = 5077a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{507760t100.pdf}

@@

with $\alpha_p= e^{i\theta_p}$ and  $\beta_p= e^{-i\theta_p}$
and
\begin{equation}
\theta_p \in [0, \pi]).
\end{equation}

@@

\begin{equation}
H_N(X): = \ \sum_{n=1}^Nc_n(H)U_n(\theta)
\end{equation}

@@

Now one does have  parity information concerning the arithmetic function $n \mapsto r_E(n)$. For a detailed study of the root numbers of $l$-functions of symmetric powers of an elliptic curve, consult \cite{DMW}.
 For $n \ge 1$ let $ \nu_E(n) \in \{0,1\}$ be (zero or one) such that  $ \nu_E(n) \equiv r_E(n)$ modulo $2$. Let $s_E(n)$ be the non-negative integer such that:
 $$r_E(n) = \nu_E(n) + 2s_E(n)$$  (for $n\ge 3$, odd).
Thus if the multiplicities of order of vanishing at the central point $s=1/2$ of the odd symmetric $n$-th power $L$-functions attached to $E$ (for $n \ge 3)$ was never greater than  $1$, and hence entirely dictated by parity, then the conjectured mean, ${\mathcal E}(S_N)$, would be equal to
\begin{equation}\label{min}
{\mathcal T}_E^{\{N\}})\ := \ {\frac{8}{3\pi}}(1-2r_E) + {\frac{2}{\pi}} \sum_{k=1}^{N}  (-1)^{k+1}\big[{\frac{1}{2k+1}} + {\frac{1}{2k+3}}\big]\big(2\nu_E(2k+1)-1\big).
\end{equation}

@@

   The standard conjecture (that we have been making all along) is GRH. But here, one includes the further conjecture (given in Sarnak's letter, and the article of Fiorilli) that the the set of nontrivial complex zeroes of the relevant $L$-function $L(E,s)$ with positive imaginary part  is a set of complex numbers that are {\it linearly independent} over ${\Q}$. Such a conjecture Rubenstein and Sarnak refer to in \cite{R-S} as the {it Grand Simplicity Hypothesis} (GSH).  Fiorilli calls his version of it  {\bf Hypothesis LI(E)}.  For recent, somewhat related, work on such linear independence questions, see cite{M-N}.  Fiorilli, following the work of Sarnak,  proves:

@@

where--following the format of explicit formulas discussed above---we view \begin{itemize} \item  the term on the LHS of the above equation as our `sum of local data'; \item  the first term on the {\rm RHS},---i.e.,  $1$---as the `{\it global term}' corresponding to the pole of $\zeta(s)$ at $s=1$ with residue $1$; it is the {\it mean} of the {\rm LHS}, our sum of local data; \item  the second term on the {\rm RHS}, $\sum_{|\gamma| \le T}{\frac{X^{i\gamma}}{({\frac{1}{2}}+i\gamma){\sqrt X}}}$ as  a cutoff at $T$ of  the `{\it oscillatory term}' while \item  the third term, $C(X,T)$ is a cutoff at $T$ and at $X$ of the `{\it easy\ error term}.'  It converges to zero if the limits are taken in the order $$\lim_{X\to \infty}\lim_{T\to \infty}C(X,T).$$ This $C(X,T)$ has the following shape:

@@

Here, $\langle X \rangle$ is the distance between $X$ and the nearest prime power, and with all this, the $<<$ would still need explicitation---even if that word is non-standard.  This result  and its proof is given, for example, as Theorem 12.5 in \cite{MV}.

@@

    \includegraphics[width=0.9\textwidth]{even_smoother-rank0-4million.png}
  \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}
   \vskip20pt
    \includegraphics[width=0.9\textwidth]{even_smoother-rank0-4million.png}
%  \includegraphics[width=0.9\textwidth]{even_smoother-rank0-4million.png}{0.8}~\label{s11}

@@

     \includegraphics[width=0.9\textwidth]{even_smoother-rank1-4million.png}{1.2}~\label{s37}
%

@@

     \includegraphics[width=0.9\textwidth]{even_smoother-rank2-4million.png}{1.2}~\label{s389}

@@

     \includegraphics[width=0.9\textwidth]{even_smoother-rank3-4million.png}{0.9}~\label{s5077}

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-rank4-4million.png}{0.9}~\label{sr4}

@@

     % \includegraphics[width=0.9\textwidth]{smooth-prime_race-rank5-4million.png}{1.2}~\label{sr5}
 \newpage

@@

     \includegraphics[width=0.9\textwidth]{even_smoother-rank6-4million.png}{1.2}~\label{sr6}
         \section{The medium-rare data:\   ${\mathcal D}_E(X)$}
       \vskip40pt

@@

  \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}
   \vskip20pt
   \includegraphics[width=0.9\textwidth]{illustsmooth-11}{.8}~\label{s11}
  \

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-37a-4million.png}{1.2}~\label{s37}
%

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-389a-4million.png}{0.9}~\label{s389}

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-5077a-4million}{0.9}~\label{s5077}

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-rank5-4million.png}{0.9}~\label{sr5}

@@

     \includegraphics[width=0.9\textwidth]{smooth-prime_race-rank6-4million.png}{1.2}~\label{sr6}
      \newpage
\section{The raw data: $\Delta_E(X)$}
\vskip40pt

@@

 \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}~ \includegraphics[width=0.9\textwidth]{normalized_straight-prime_race-11a-4million.png}{.8}~\label{nr11}

@@

     \includegraphics[width=0.9\textwidth]{normalized_straight-prime_race-37a-4million.png}{.8}~\label{nr37}
%

@@

     \includegraphics[width=0.9\textwidth]{normalized_straight-prime_race-389a-4million.png}{.8}~\label{nr389}

@@

     \includegraphics[width=0.9\textwidth]{normalized_straight-prime_race-5077a-4million.png}{.8}~\label{nr389}

>>>>

The  term `Explicit Formula' as in our title refers to the genre of formula that expresses {\it arithmetically interesting quantities} in terms of the zeroes of  a related zeta-function or $L$-function.  The first such formula appears towards the end of Bernhard Riemann's short and great paper  ``\"{U}ber die Anzahl der Primzahlen unter einer
gegebenen Gr\"{o}sse," published in the
Monatsberichte der Berliner Akademie,
November 1859.  Subsequently there have been many versions of Explicit Formulas, one such notable variant  proved by Hans von  Mangoldt, in his 1895 article  "Zu Riemann's Abhandlung ``\"{U}ber die Anzahl der Primzahlen unter einer gegebenen Gr\"{o}sse"", Journal f\"{u}r die reine und angewandte Mathematik.  We devote this expository  article in honor of Riemann to a discussion of formulas of this type. Our main aim is to formulate some open questions related to the application of the Explicit Formula to the arithmetic of elliptic curves.

  The {\it arithmetically interesting quantity} that appears in most of the `Explicit Formulas' of analytic number theory is often  a partial sum $F(X)\cdot\sum_{p<X}G(p)$ of locally defined quantities $G(p)$ attached to prime numbers $p$, summed to up to some cutoff value, $p<X$  (and normalized for convenience by an elementary factor $F(X)$). The formula expresses this function of the variable $X$, $F(X)\cdot\sum_{p<X}G(p)$, as a {\it dominant term}, plus an easily controlled error term, plus an interesting third term that might be called the {\it oscillatory term} that, in most cases, is only conjecturally controllable.

@@

  Specifically we will  focus on {\it issues of bias}  following the classical ``Explicit Formula,'' and the work of:
Sarnak, \ Granville,\  Rubenstein,\, Martin-Watkins,\ Fiorilli,\ Bober,\ Conrey-Snaith, and others.
 The example-problem we consider is related to the question---given an elliptic curve over the rational numbers and letting  $p$ range through prime numbers---of how often  $p+1$ is an {\it over-count} or an {\it under-count} for the number of rational points on the curve modulo $p$? The rough answer is 50/50, but there can be a `bias' similar to the classical Chebyshev bias  \cite{R-S}. For such finer statistics one resorts to `Explicit Formulas.'  Here, computation can even outstrip theory in that people have algorithms to make such computations whether or not the holomorphicity of the $L$-functions in question have been proved. Computations have depended on the work of many people.

 We have used this text to provide material for {\it unusually short lectures}. %Specifically, the notes were initially written for each of our 20 minute lectures at the San Diego Conference, as we have already mentioned, but also
 E.g., one of us replayed it all{\footnote{ Well, not quite all$\dots$}} in a 6 minute and 40 second lecture at Harvard. We say this to make it clear that this article is meant to be  somewhat light reading; it contains no new results, as we have mentioned, and no proofs of anything. But we feel that the issues here are a good source of student projects, and that our text might be  a companion to any reader of the beautiful expository articles on classical Chebyshev bias, such as \cite{R-S}, or on the biases we discuss, as in \cite{S}.

@@

 Happily, the other of us (W.S.) has produced relevant computations that do exactly that (for applications of the Explicit Formula to certain problems of current interest to both of us).% This, then, is a phenomenological talk, with (at least the beginning of)  a corpus  of graphs that offer some  illustration of the effectiveness---or non-effectiveness---of the explicit formula in the specific instances of interest to us.

@@

  \section{ The RHS of our Explicit Formulas}
   Here are some brief comments on each of the three  `terms' (i.e., {\it Global data}, {\it  Easy error term}, and {\it  Oscillatory term})  on the RHS of the Explicit  Formula for the well-done variant (see above) for the elliptic curve $E$:

@@

  \begin{conjecture}  $ S_E(X)\ \  {\stackrel{?}{=}}\ \  o(\log X).$\end{conjecture}   The analogous oscillatory term for the classical Riemann zeta function  has an extensive literature. See, for example \cite{G}, \cite{Fu} and the bibliography there. Here are some pictures to convey a sense of how our $S_E(X,T)$ behaves, at least in the currently computable range which (roughly) allows $T$ to be only as high as  $10^4$.
  \newpage
  \centerline{ $E = 11a$}
 \vskip10pt
  \includegraphics[width=0.9\textwidth]{plots/OSC11.pdf}

  \centerline{ $E = 37a$}
  \includegraphics[width=0.9\textwidth]{plots/OSC37.pdf}
%\vskip20pt
 % \includegraphics[width=1.0\textwidth]{plots/OSC.pdf}
   \vskip20pt
  \centerline{ $E = 389a$}
\vskip20pt
  \includegraphics[width=0.9\textwidth]{plots/OSC389.pdf}

@@

Let ${\R}_{>0}$ be the multiplicative group of positive real numbers, and ${\R}$ the additive group of reals. For $I \subset {\R}_{>0}$ a Haar measurable set, let $|I|$ denote a Haar measure.  Let ${\S}:{\R}_{>0} \to {\R}$ be a real-valued Lebesgue-integrable function.  Fixing $I \subset {\R}_{>0}$ a subset  of finite measure,  for every measurable subset $J\subset {\R}$, form the probability measure on ${\R}$
$$J\mapsto \mu_{{\S},I}(J): = {\frac{|I\cap {\S}^{-1}(J)|}{|I|}}.$$  So, $\mu_{{\S},I}(J)$ is the {\it probability that the function ${\S}$ achieves a value in the range $J$ over the gamut of arguments in $I$.}   Say that ${\S}$ has a {\bf normal distribution of values} if, for $X > 0$ setting $I_X= (0,X]$, the limit  $$\mu_{{\S}}:= \lim_{X \to \infty}\mu_{{\S},I_X}$$ exists. These  definitions are particularly relevant to the oscillatory terms ${\S}(X):= S_E(X)$ that we are currently studying. The data seems to indicate convergence to a limiting distribution (the  mean  value being $0$) with a strikingly small (variance, or equivalently: strikingly small) standard deviation of values.
\vskip20pt
Here, then, are some pictures  of what seems to be data 'converging' to a limiting  distribution $\mu_E$ of the values of the oscillatory terms $S_E(X)$ for a few elliptic curves $E$:
  \vskip10pt
  \centerline{ $E = 11a$}
 \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/bite11.pdf}
    \vskip10pt
  \centerline{ $E = 37a$}
  \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/bite37.pdf}
    \vskip10pt
       The red curve is the normal distribution with mean $0$ and standard deviation given by that of the data.
   \vskip10pt
    {\bf  Note: } Conditional on the conjecture $LI(E)$ (see section \ref{Fi} below) $\mu_E$ exists (see \ref{S}).

     It is interesting  to compare $\mu_E$ to the limiting distributions connected to the bias of nonresidues to residues mod $q$, as in \cite{R-S}. There one has the added feature that these limiting distributions themselves tend to the normal distribution as the modulus q tends to infinity.
   % \includegraphics[width=0.9\textwidth]{plots/bite389.pdf}
 %\vskip30pt
  \centerline{ $E = 389a$}
  \vskip10pt
\hskip100pt  \includegraphics[width=0.6\textwidth]{plots/bite389.pdf}

@@

 \centerline{ $E = 11a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/11a40t60.pdf}
    \newpage
  \centerline{ $E = 11a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/11a60t100.pdf}

   \vskip10pt
   \centerline{ $E = 11a$  between $990$ and $1010$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/11a990t1010.pdf}

@@

   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/37a40t60.pdf}

   \centerline{ $E = 37a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/37a60t100.pdf}
 \newpage
  \centerline{ $E = 37a$  between $990$ and $1010$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/37a990t1010.pdf}
 \newpage
  \centerline{ $E = 389a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/389a40t60.pdf}

   \centerline{ $E = 389a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/389a60t100.pdf}
\newpage
  \centerline{ $E = 5077a$  between $40$ and $60$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/5077a40t60.pdf}

   \centerline{ $E = 5077a$  between $60$ and $100$}
   \vskip10pt
 \hskip100pt \includegraphics[width=0.6\textwidth]{plots/507760t100.pdf}

@@

with $\alpha_p= e^{i\theta_p}$ and  $\beta_p= e^{-i\theta_p}$
and
\begin{equation}
  \theta_p \in [0, \pi].
\end{equation}

@@

\begin{equation}
H_N(\theta): = \ \sum_{n=1}^Nc_n(H)U_n(\theta)
\end{equation}

@@

Now one does have  parity information concerning the arithmetic function $n \mapsto r_E(n)$. For a detailed study of the root numbers of $L$-functions of symmetric powers of an elliptic curve, consult \cite{DMW}.
 For $n \ge 1$ let $ \nu_E(n) \in \{0,1\}$ be (zero or one) such that  $ \nu_E(n) \equiv r_E(n)$ modulo $2$. Let $s_E(n)$ be the non-negative integer such that:
 $$r_E(n) = \nu_E(n) + 2s_E(n)$$  (for $n\ge 3$, odd).
Thus if the multiplicity of order of vanishing at the central point $s=1/2$ of the odd symmetric $n$-th power $L$-functions attached to $E$ (for $n \ge 3)$ were never greater than  $1$, and hence entirely dictated by parity, then the conjectured mean, ${\mathcal E}(S_N)$, would be equal to
\begin{equation}\label{min}
{\mathcal T}_E^{\{N\}}\ := \ {\frac{8}{3\pi}}(1-2r_E) + {\frac{2}{\pi}} \sum_{k=1}^{N}  (-1)^{k+1}\big[{\frac{1}{2k+1}} + {\frac{1}{2k+3}}\big]\big(2\nu_E(2k+1)-1\big).
\end{equation}

@@

   The standard conjecture (that we have been making all along) is GRH. But here, one includes the further conjecture (given in Sarnak's letter, and the article of Fiorilli) that the the set of nontrivial complex zeroes of the relevant $L$-function $L(E,s)$ with positive imaginary part  is a set of complex numbers that are {\it linearly independent} over ${\Q}$. Such a conjecture Rubenstein and Sarnak refer to in \cite{R-S} as the {\it Grand Simplicity Hypothesis} (GSH).  Fiorilli calls his version of it  {\it Hypothesis LI(E)}.  For recent, somewhat related, work on such linear independence questions, see \cite{M-N}.   Fiorilli, following the work of Sarnak,  proves:

@@

where---following the format of explicit formulas discussed above---we view \begin{itemize} \item  the term on the LHS of the above equation as our `sum of local data'; \item  the first term on the {\rm RHS},---i.e.,  $1$---as the `{\it global term}' corresponding to the pole of $\zeta(s)$ at $s=1$ with residue $1$; it is the {\it mean} of the {\rm LHS}, our sum of local data; \item  the second term on the {\rm RHS}, $\sum_{|\gamma| \le T}{\frac{X^{i\gamma}}{({\frac{1}{2}}+i\gamma){\sqrt X}}}$ as  a cutoff at $T$ of  the `{\it oscillatory term}' while \item  the third term, $C(X,T)$ is a cutoff at $T$ and at $X$ of the `{\it easy\ error term}.'  It converges to zero if the limits are taken in the order $$\lim_{X\to \infty}\lim_{T\to \infty}C(X,T).$$ This $C(X,T)$ has the following shape:

@@

Here, $\langle X \rangle$ is the distance between $X$ and the nearest prime power, and with all this, the $<<$ would still need ``explicitation''---even if that word is non-standard.  This result  and its proof is given, for example, as Theorem 12.5 in \cite{MV}.

@@

    \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank0-4million.png}
  \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}
   \vskip20pt
    \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank0-4million.png}
%  \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank0-4million.png}{0.8}~\label{s11}

@@

     \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank1-4million.png}{1.2}~\label{s37}
%

@@

     \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank2-4million.png}{1.2}~\label{s389}

@@

     \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank3-4million.png}{0.9}~\label{s5077}

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-rank4-4million.png}{0.9}~\label{sr4}

@@

     % \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-rank5-4million.png}{1.2}~\label{sr5}
 \newpage

@@

     \includegraphics[width=0.9\textwidth]{plots/even_smoother-rank6-4million.png}{1.2}~\label{sr6}
         \section{The medium-rare data:\   ${\mathcal D}_E(X)$}
       \vskip40pt

@@

  \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}
   \vskip20pt
   \includegraphics[width=0.9\textwidth]{plots/illustsmooth-11}{.8}~\label{s11}
  \

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-37a-4million.png}{1.2}~\label{s37}
%

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-389a-4million.png}{0.9}~\label{s389}

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-5077a-4million}{0.9}~\label{s5077}

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-rank5-4million.png}{0.9}~\label{sr5}

@@

     \includegraphics[width=0.9\textwidth]{plots/smooth-prime_race-rank6-4million.png}{1.2}~\label{sr6}
      \newpage
\section{The raw data: $\Delta_E(X)$}
\vskip40pt

@@

 \centerline{\bf Rank $r=0$:\ \ \  ${\mathcal E}=$11A.}~ \includegraphics[width=0.9\textwidth]{plots/normalized_straight-prime_race-11a-4million.png}{.8}~\label{nr11}

@@

     \includegraphics[width=0.9\textwidth]{plots/normalized_straight-prime_race-37a-4million.png}{.8}~\label{nr37}
%

@@

     \includegraphics[width=0.9\textwidth]{plots/normalized_straight-prime_race-389a-4million.png}{.8}~\label{nr389}

@@

     \includegraphics[width=0.9\textwidth]{plots/normalized_straight-prime_race-5077a-4million.png}{.8}~\label{nr389}
